### 2017-08-21 pea_paper_docs/tex/pea.tex
fixed URL in README.md

\documentclass[10pt, onecolumn]{article}
\usepackage[margin=1in]{geometry}
\usepackage{lmodern}% http://ctan.org/pkg/lm
\usepackage{authblk} % adds affiliations

\usepackage[utf8x]{inputenc}
\usepackage{nameref}
\usepackage[right]{lineno}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage[numbers,super]{natbib}
\usepackage{changepage}

% adjust caption style
\usepackage[aboveskip=1pt,labelfont=bf,
            labelsep=period,singlelinecheck=off]{caption}
% remove brackets from references
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% headrule, footrule and page numbers
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
% \fancyheadoffset[L]{2.25in}
% \fancyfootoffset[L]{2.25in}
% use \textcolor{color}{text} for colored text (e.g. highlight to-do areas)
\usepackage{color}
% define custom colors (this one is for figure captions)
\definecolor{Gray}{gray}{.25}
% this is required to include graphics
\usepackage{graphicx}
% set the figures directory
\graphicspath{{./figs/}}
% use if you want to put caption to the side of the figure - see example in text
\usepackage{sidecap}
% \usepackage[urlcolor  = blue]{hyperref}

% hyperurls packages:
\usepackage{xcolor}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}

% use for have text wrap around figures
\usepackage{wrapfig}
\usepackage[pscoord]{eso-pic}
\usepackage[fulladjust]{marginnote}
\reversemarginpar{}

% new commands
\newcommand{\cel}{\emph{C.~elegans}}
\newcommand{\fog}{\emph{\mbox{fog-2}}}
\newcommand{\ecol}{\emph{E.~coli}}
\newcommand{\hobesity}{957}
\newcommand{\wobesity}{614}
\newcommand{\hlupus}{283}
\newcommand{\wlupus}{135}
\newcommand{\harthritis}{309}
\newcommand{\warthritis}{124}

\newcommand{\qval}[1]{\ensuremath{q<10^{-#1}}}

% more space between rows
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\title{
  \Large
  \textbf{
  % Phenotype Enrichment Discovers Phenologs for Disease Modeling in \cel{}
  Phenotype and gene ontology enrichment as guides for disease modeling
  in \cel{}
          }
}

\author[1,2]{David Angeles-Albores}
\author[1]{Raymond Y. N. Lee}
\author[1]{Juancarlos Chan}
\author[1,2,*]{Paul W. Sterberg}
\affil[1]{Division of Biology and Biological Engineering, Caltech,
Pasadena, CA, 91125, USA}
\affil[2]{Howard Hughes Medical Institute, Caltech, Pasadena, CA, 91125, USA}
\renewcommand\Affilfont{\itshape\small{}}

\begin{document}

% \vspace*{0.35in}

% title goes here:
% title
\maketitle

\textbf{1} Department of Biology and Biological Engineering,
and Howard Hughes Medical Institute, Caltech, Pasadena, CA, 91125, USA\\
\textbf{*} Corresponding author. Contact:pws@caltech.edu

\section*{Abstract}
\textbf{
Genome-wide experiments have the capacity to generate massive amounts of unbiased
data about an organism. To interpret these data, dimensionality reduction
techniques are required. One approach is to annotate genes using unified,
unambiguous vocabularies and to test experimental datasets for term enrichment
using probabilistic methods.
Although gene, phenotype and anatomy ontologies exist for \cel{}, no one
software offers enrichment analyses of all three ontologies.
Here, we present the WormBase Enrichment Suite, which offers users the
ability to test all nematode ontologies simultaneously. We show that
the WormBase Enrichment Suite provides valuable insight into different
biological problems, and that phenotype enrichment analysis (PEA) can
help researchers identify disease phenologs, phenotypes that are homologous across
species, which can inform disease modeling in \cel{}. The WormBase Enrichment Suite
analysis can also shed light on RNA-seq datasets by showing what
molecular functions are enriched, which phenotypes these functions are implicated
in and what tissues are overrepresented in the dataset. Finally, we explore the
phenotype-anatomy relationship, showing that a small subset of highly specific
tissues are disproportionately likely to cause an Egl phenotype, but inferring
tissue expression from an egg-laying (Egl) phenotype is limited to the largest tissues.
}

\vspace{10mm}

% now start line numbers
\nolinenumbers{}

\section*{Introduction}
The last decade has seen an explosion of techniques capable of genome-wide
measurements. Some examples of genome-wide tools include
RNA-seq~\cite{Mortazavi2008} to measure gene expression or
CHIP-seq~\cite{Johnson2007} to measure protein binding to chromatin. These tools
are capable of generating large quantities of data. Understanding these data,
and generating hypotheses from them remains challenging. A common approach used
to understand these datasets is to reduce the dimensionality of the data via
enrichment analyses of ontologies~\cite{TheGeneOntologyConsortium2000a}, which
helps researchers understand what terms are overrepresented beyond random
levels. By analyzing overrepresented terms in aggregate, researchers can better
understand what biological processes were most affected in a given experiment,
and form hypotheses about what is happening~\cite{Rhee2008}. This approach is
limited by what ontologies can be tested for enrichment. The best-known ontologies
for biological research are the Gene Ontologies (GO), which provide a controlled
language to describe the biological, cellular and molecular aspects and functions of
genes~\cite{TheGeneOntologyConsortium2000a}. In \cel{}, gene, tissue and
phenotype ontologies exist with which to describe \cel{} anatomy and phenotypes
respectively~\cite{Schindelman2011,Lee2003}. These latter ontologies are developed by
professional curators at WormBase, which is a repository of \cel{}
data~\cite{Howe2016}. However, enrichment tools currently only exist for gene and tissue
ontologies in the community today (see for
example~\cite{Chikina2009,Mi2013,Angeles-Albores2016}). Another limitation is
that tissue enrichment testing is not offered on the same websites as GO
enrichment testing, which requires users to test their data on different
websites that may or may not use different methodologies to detect enrichment.

Another way to use enrichment tools is for evolutionary comparison purposes.
In molecular biology it is often useful to know when a gene is homologous
between two species---that is to say, common by descent---because knowledge of
homology often brings with it knowledge of function. Indeed, many
important gene regulatory networks (GRNs) are conserved between organisms as
highly diverged as nematodes and humans (for example, see~\cite{Sternberg1998}).
While genes and GRNs may be conserved between species, their outputs often differ.
For example, the gene Pax6 (Eyeless in \emph{Drosophila melanogaster})
is involved in eye formation in humans and fruit-flies~\cite{Quiring1994}. Although
nematodes have conserved this gene, they do not have eyes~\cite{Zhang1995,Chisholm1995}.
The concept of a phenolog has been put forward to explain relationships between
phenotypes that have the same underlying genetic
regulatory network~\cite{McGary2010,Lehner2013}.
Formally, two phenotypes are phenologs of each other if the orthologs of the
genes that cause a phenotype in an organism cause a second phenotype in another.

To study a clinically relevant disease in a non-human, an appropriate model has
to be established. A straightforward method towards establishing a disease model
in \cel{} is to link a disease to a causal gene, then to identify the homologous
gene in \cel{} and then to study the function of the genetic homolog to
extrapolate back to humans. However, this method relies on the existence of
known disease genes and requires that the homolog have a phenotype that can be
reliably identified and studied. A fundamentally different way to establish a
disease model in \cel{} would be to identify the phenologs of the disease to be
studied in \cel{} by identifying disease-associated human genes in an unbiased
manner through genome-wide association studies (GWAS) and identified candidate
homolog genes in \cel{}. The orthologs can be used to identify \cel{} disease
phenologs, which can in turn be used as the basis for screens to identify genes
that are associated with that phenolog. Approaches similar to this have been
successfully used in the past to make non-obvious links between phenotypes in
different species~\cite{McGary2010}.

The concept of a phenolog can also be useful when applied within a species. In
\cel{}, not all phenotypes are equally easy to study. Although genome-wide
measurements can help elucidate the genetic network underlying a phenotype,
devising screens to test which genes are functionally important can be
difficult. A common strategy to study phenotypes that are difficult to screen is
to select an easier-to-screen phenolog, and to test positive hits for the true
phenotype of interest afterwards. For example, candidate genes to extend the
\cel{} lifespan can be first screened for using heat shock survival genes
involved in \cel{} aging sensitivity can be identified using stress
assays~\cite{Kim2007a,Mehta2009}. Currently, selection of screening phenotypes
is performed based on researcher experience. By formalizing phenotype enrichment
analysis as a tool with which to analyze gene sets, researchers should be able
to formally establish phenologs, which has consequences for screen design. %
eating defects in an ablated strain

An additional problem with genome-wide queries of \cel{} states (be they
developmental, such as L1, L2, dauer; behavioral states such as awake versus
asleep; or other) is that they do not always have a straightforward
interpretation in terms of phenotypes. In these situations, researchers must
rely on intuition to select a phenotype for which to screen. As a result, many
hits may go unexplored that would prove fruitful. The question of how to design
a screen that is maximally informative is an important question that has so far
not been addressed within this community.

To facilitate understanding of large datasets, we have developed the WormBase
Enrichment Suite (URL:\
\url{http://www.wormbase.org/tools/enrichment/tea/tea.cgi}) that allows users to
rapidly perform phenotype, tissue and gene ontology enrichment analyses (PEA,
TEA and GEA respectively) on curated \cel{} ontologies using the same
methodology for each one. As examples of use cases, we applied our tools towards
the unbiased discovery of phenologs of multigenic, complex diseases including
systemic lupus erythematosus, obesity and obesity-related traits, and rheumatoid
arthritis by using genes associated with these diseases via genome-wide
association studies. We illustrate the utility of the complete enrichment suite
for finding new relationships in complex data by analyzing a ciliary neuron
transcriptome~\cite{Wang2015}. Finally, we show that the dictionaries generated
for these enrichment analyses can help elucidate the contributions of specific
tissues to specific phenotypes.

\section*{Methods}
\subsection*{Implementing the enrichment analyses}
All scripts were implemented in Python 3.5~\cite{Rossum2011}. We used
pandas~\cite{McKinney2011} and scipy~\cite{Oliphant2007} to write the
statistical testing framework. Matplotlib~\cite{Hunter2007} and
Seaborn~\cite{Waskom} libraries are used to generate all plots. Testing was
performed using the WormBase version WS256. The WormBase Enrichment suite can be
installed using pip via the command:

\texttt{pip install tissue\_enrichment\_analysis}

Enrichment analyses are performed as described in~\cite{Angeles-Albores2016}.
Briefly, a static dictionary is constructed for each ontology using the most
recent WormBase release annotation information. This dictionary is generated
using a trimming algorithm described in~\cite{Angeles-Albores2016} to remove
terms that are too general and uninformative as well as terms that are very
specific but do not have many annotations. When the user provides a set of genes
to test against, we use a hypergeometric model to test whether a given term
within an ontology is occurring at a greater frequency than expected by random
chance.

\subsection*{Human disease phenolog identification}
We used the GWAS EBI-NHGRI catalog~\cite{MacArthur2016} to extract information
on all genome-wide association studies deposited there. We only selected traits
that had $>300$ associated genes. We identified 24 traits that met our criteria.
Next, we used DIOPT~\cite{Hu2011} to identify candidate orthologs for the genes
associated with these traits. Briefly, DIOPT combines a large number of methods
for identifying orthologs and returns homolog candidates associated with a
compound score. Depending on the score, orthologs can be considered `high',
`moderate' or `low' rank, reflecting confidence in the homology. Many-to-one and
one-to-many homology relationships are allowed in DIOPT, reflecting a mixture of
uncertainty and family expansion/reduction. For our study, we only accepted
homolog candidates with `high' or `moderate' scores and we did not insist on a
one-to-one relationship between genes.

\begin{figure}[htbp]
  \renewcommand{\familydefault}{\sfdefault}\normalfont{}
  \centering
  \includegraphics[width=.5\linewidth]{gwas-design.pdf}
  \caption{Experimental design for human-nematode phenolog identification.
           We used GWAS candidates from the EBI-NHGRI catalog to identify
           disease associated genes, then used DIOPT to identify candidate
           orthologs in \cel{}. Orthologs were used to run phenotype, gene
           and tissue ontology enrichment analyses to identify disease
           phenologs.}
\label{fig:gwas}
\end{figure}

After we identified worm orthologs for each trait, we reassessed how many traits
still had $>100$ gene candidates, and dropped all traits that had less than this
for our analysis. We identified 18 traits that met this criteria. The gene lists
for each of these 18 traits were then analyzed for gene, tissue and phenotype
enrichment (see~Fig.\ref{fig:gwas}). Tissue enrichment was performed using the
WormBase Tissue Enrichment Analysis (TEA) tool~\cite{Angeles-Albores2016}.

\section*{Results}
\subsection*{Developing the WormBase enrichment suite}
% TODO add numbers instead of XXX
We developed the dictionaries for PEA and GEA using the same procedure as was
used for TEA~\cite{Angeles-Albores2016}. We generated a dictionary that included
terms with at least 50 annotating genes or more and had a similarity threshold
of 0.95 for PEA (the total number of terms in the dictionary was 251, annotated
by 9,169 genes for the version WS256);\@ and we generated a dictionary that
included terms with at least 50 annotating genes or more and had a similarity
threshold of 0.95 for GEA (the total number of terms in this dictionary is 271,
annotated by 14,636 genes for the version WS256). \@ Next, we benchmarked the
dictionaries on the same gene sets as TEA and obtained enrichment of all the
expected categories~\cite{Gaudet2004a, Spencer2011, Cinar2005, Watson2008a,
Pauli2006, Portman2004, Fox2007, Smith2010}. For example, on a gene set enriched
for embryonic muscle genes~\cite{Watson2008a}, the top two enriched phenotype
terms by q-value were `muscle system morphology variant' and `body wall muscle
thick filament variant'; the top two enriched GO terms were `myofibril' and
`striated muscle dense body'. For all the benchmarking results, see
supplementary information. Having generated and validated our dictionaries, we
proceeded to identify phenologs for several common human diseases.

\subsection*{Applying the WormBase enrichment suite}
To discover phenologs, we first needed to identify genes that contribute to a
disease in an unbiased manner. One way to discover gene associations in an
unbiased manner is to perform GWSA in human populations. Therefore, we used the
GWAS NHGRI-EBI Catalog~\cite{MacArthur2016} to identify genes associated with
human diseases. We found the best nematode candidate orthologs for these genes
using DIOPT~\cite{Hu2011} and applied our enrichment suite to each of these gene
regulatory networks.

\subsection*{Systemic lupus erythematosus}
Systemic lupus is an autoimmune disease that is believed to be polygenic in
nature~\cite{Mohan2015}. It mainly affects women and is characterized by painful
and swollen joints, hair loss, and fatigue~\cite{Lisnevskaia2014}. Since worms
do not have a cellular immune system, we were interested in what phenologs
corresponded to this disorder in \cel{}. To establish phenolog candidates, we
obtained \hlupus{} genes associated with the disease via GWAS studies, and found
\wlupus{} homolog candidates in \cel{}.

Lupus-associated orthologs were reasonably well annotated. Slightly more than
half of the genes had at least one phenotype annotation ($76/\wlupus{}$) and
almost all genes were annotated to at least one tissue or gene ontology term
($104/\wlupus{}$ and $115/\wlupus$ genes respectively). We found that
Lupus-associated orthologs were enriched in `aneuploidy' (7 genes, \qval{1}) and
`meiotic chromosome segregation' (8 genes, \qval{1}). `Cell fate transformation'
(6 genes, \qval{1}), and `excess intestinal cells' (5 genes, \qval{1}) were also
overrepresented, as was `male tail morphology' (6 genes, \qval{1}). Finally, the
phenotype `nonsense mRNA accumulation' was also enriched (5 genes, \qval{1})
(see Fig.~\ref{fig:lupus}).

\begin{figure}[htbp]
  \renewcommand{\familydefault}{\sfdefault}\normalfont{}
  \centering
  \includegraphics[width=.5\linewidth]{systemic-lupus.pdf}
  \caption{phenolog identification for systemic lupus erythematosus.
           \textbf{A} Phenotype Enrichment Analysis. \textbf{B} GO Enrichment
           Analysis. \textbf{C} Lupus in \cel{} may be best represented by
           a combination of three phenotypes: Cell proliferation possibly
           accompanied by aneuploidy; cell fate transformations that may lead
           to dysmorphias; and a molecular phenotype involving impairment of
           the Nonsense-Mediated Decay (NMD) pathway.}
\label{fig:lupus}
\end{figure}

TEA suggested that the `excretory duct cell' (5 genes, \qval{2}) and the
`posterior gonad arm' are overrepresented in this dataset. We also found that
the Pn.p cells P3.p through P8.p were enriched in this dataset (5 genes,
\qval{1}). GO enrichment pointed at `modification-dependent macromolecule
catabolic process' (23 genes, \qval{15}) as a molecular function that
characterizes this dataset. However, this GO term was enriched only due to a
single gene family, the \emph{skr} gene family. Almost the entire \emph{skr}
family was considered a candidate homolog to the SKP1 human gene, making the GO
enrichment suspect.

Enrichment of the terms for `aneuploidy', `meiotic chromosome segregation', and
`excess intestinal cells' were largely driven by the same gene group, which
includes \emph{cki-1}, and several \emph{skr} genes. On the other hand, `cell
fate transformation' and `male tail morphology' reflected the involvement of
developmental genes \emph{let-23}, and \emph{lin-12} among others. The term
`nonsense mRNA accumulation' was the result of \emph{pept-3}, \emph{smg-7},
\emph{tsr-1}, \emph{dhcr-7} and \emph{F08B4.7}. Therefore, we conclude that
systemic lupus erythematosus is potentially represented by a combination of
three phenotypes in \cel{}: A cell proliferation phenotype (either increased or
decreased), probably marked by increased aneuploidy; a developmental phenotype
involving cell fate transformation and leading to dysmorphias; and a molecular
phenotype involving impairment of the nonsense-mediated decay pathway. % TODO:
Is the text below true??? The results from the tissue enrichment analysis
highlighted three tissues that are particularly sensitive to \emph{lin}
mutations (the gonad, the excretory duct cell and the vulval precursor cells),
and the gonad arms undergo large quantities of nuclear proliferation.

\subsection*{Rheumatoid arthritis}
Rheumatoid arthritis is an auto-immune disease that is characterized by swollen
and painful joints that progressively deteriorate~\cite{Smolen2016}. Unlike
lupus, rheumatoid arthritis is not life-threatening, and comorbidity between
rheumatoid arthritis and lupus has not been described in past comorbidity
studies~\cite{Dougados2013}, suggesting that they may have at least partially
distinct genetic causes. We found \harthritis{} genes associated with rheumatoid
arthritis, for which we found \warthritis{} worm homolog candidates.

% TODO: Is no tissue enrichment observed bc poor annotations, or because
% all tissues randomly represented?
The only phenotype that was enriched for these orthologs was `short' (10 genes,
\qval{4}), even though 64 orthologs were associated with at least one phenotype
term. No tissue was enriched in this dataset. Because 82 genes are annotated to
have expression in at least one tissue, the lack of enrichment does not reflect
ignorance about the sites of expression of these genes. GEA showed that enriched
molecular functions for these genes include `collagen trimer' (22 genes,
\qval{15}). However, this term was enriched as the result of degeneracy in the
homolog candidates for the SFTPD gene. Other terms included `glycosylation' (10
genes, \qval{4}) and `Golgi apparatus' (11 genes, \qval{3}), but these terms
were enriched as the result of degenerate homolog candidates for the human gene
B3GNT7 which encodes a beta-1,3-N-acetylgalactosaminyltransferase.

The `short' phenotype was the result of the \emph{cat-4}, \emph{dpy-7},
\emph{rnt-1}, \emph{sem-4}, \emph{unc-116}, \emph{ocrl-1} and some genes in the
\emph{fat} family. Although these genes are bound by a common phenotype, any
genetic relationships between these genes are not immediately clear. Some genes,
like \emph{sem-4} and \emph{rnt-1} are likely transcription factors with roles
in development (including hypodermal development)~\cite{Desai1988,Ji2004}.
Others are molecular motors (\emph{unc-116}) that are broadly expressed
throughout the body of \cel{}. Yet others have known roles in neuron and muscle
function, such as \emph{ocrl-1} and \emph{cat-4}. The `short' phenotype is a
subset of the `body length variant' phenotype. Body length in \cel{} can be
controlled via cell size or shape\cite{Wang2002,Nakano2004}; alternatively,
cuticle development can alter body shape~\cite{Cox1980,Kramer1988}; finally,
muscles can alter the effective body length due to their contraction
state~\cite{Lewis1980}.

\subsubsection*{Obesity-related traits}
Obesity-related traits is a category within the GWAS NHGRI-EBI catalog that
pools studies that have measured obesity and other traits associated with
obesity, such as heart rate, physical activity, hormone levels, body composition
and cholesterol levels. Since this category includes many parameters, we
expected there would be many phenologs. GWAS studies have identified \hobesity{}
genes associated with these traits. Using DIOPT, we found \wobesity{} orthologs
for these genes. In total, $341/\wobesity{}$ genes had at least one phenotype
annotation; $548/\wobesity{}$ had at least one gene ontology term annotation;
and $427/\wobesity{}$ had at least one tissue term annotation.

Top results for obesity-related traits included `acetylcholinesterase inhibitor
response variant' (38 genes, \qval{6}), `neurite morphology variant' (21 genes,
\qval{2}), and `thin' (31 genes, \qval{2}). Terms involving locomotion were
significantly enriched, as were terms involving body shape and food consumption
(\qval{1}). Concomitant with these phenologs was a tissue enrichment in
neuron-related terms. GO enrichment suggested that these genes are participating
in `iron ion binding' (40 genes, \qval{20}) and `tetrapyrrole binding' (37
genes, \qval{13}).

Tissue and phenotype enrichment therefore suggest that obesity-related traits
may be studied in \cel{} through neuron physiology and function, specifically
with respect to acetylcholinesterase inhibitors. Moreover, GO enrichment
implicates iron and tetrapyrrole binding as metabolic components of the
obesity-related phenologs in \cel{}.

\subsection*{Ontology enrichment as an aid for screen design}
An additional use for a tool like PEA would be as a tool to help guide and
design screens to identify genes from an RNA-seq or other genome-wide experiment
for further study. This would be particularly useful in cases when researchers
may not know what phenotype to expect, in which case PEA can guide selection of
a phenotype. Another use case is a scenario where the phenotype under study is
not easy to screen for. By finding phenologs to the phenotype of interest, the
researcher can design an easier screen for genes that affect the phenolog in
question, then re-test genes for the original phenotype of interest.

\subsubsection*{Enrichment in the ciliary neuronal transcriptome}
As an example of how ontology enrichment can improve our understanding of
transcriptomes, we selected a ciliary neuron dataset~\cite{Wang2015} and ran the
complete WormBase Enrichment Suite on it. Ciliary neurons are present in the
\cel{} male tail, but they are also present in the male cephalic sensillum and
hermaphrodites also have ciliated neurons. PEA reveals that the ciliated neuron
transcriptome is enriched for genes that are typically associated with `meiotic
chromosome segregation' (46 genes, \qval{5}), `aneuploidy' (42 genes, \qval{5})
and `spindle defective early embryos' (45 genes, \qval{2}) (see
Fig.~\ref{fig:cilia}).

\begin{figure}[htbp]
  \renewcommand{\familydefault}{\sfdefault}\normalfont{}
  \centering
  \includegraphics[width=.5\linewidth]{ciliary-transcriptome.pdf}
  \caption{PEA shows that the ciliary transcriptome is enriched in phenotypes
  related to cell division. Although this could reflect enrichment of
  microtubules and microtubule-related genes, the enrichment is at least
  partially driven by cell-cycle and DNA repair genes.}
\label{fig:cilia}
\end{figure}

In addition, TEA points at the \cel{} gonad primordium, the somatic gonad and
early embryonic cells as the sites where genes associated with ciliary neurons
are enriched. The `male distal tip cell' is a tissue that is overrepresented in
this dataset, but `distal tip cell' is not enriched. In \cel{} the hermaphrodite
distal tip cells (DTCs) and the male DTC are very similar to each other in their
biological functions (both maintain a stem cell niche in the distal gonad).
However, the male DTC is non-migratory, whereas the hermaphrodite DTCs are
migratory. Therefore, the term `male distal tip cell' may reflect cellular
aspects that are correlated with the non-migratory aspects of the DTC biology,
such as maintenance of proliferation.

Although one interpretation of the results would be that microtubule genes are
driving the enrichment of these terms, another possibility is that there are
cell-cycle genes that are driving the enrichment of these phenotypes and
tissues. Indeed, GO enrichment shows terms such as `DNA replication' (29 genes,
\qval{5}), and `purine NTP-dependent helicase activity' (15 genes, \qval{1}).
Visual inspection of list in question reveals that cell-cycle and DNA
replication/repair genes are abundant in this transcriptome and include genes
such as \emph{atm-1}, \emph{dna-2}, or \emph{hpr-17}. This analysis reveals that
the ciliary neuron transcriptome is enriched in genes associated with
microtubules, but includes genes that are thought to interact with DNA either
via repair mechanisms or cell-cycle
control~\cite{Hofmann2000,Lee2003a,Bailly2010}.

\subsection*{Deconstructing phenotype-tissue relationships}
\subsubsection*{Tissue enrichment on the Egl gene set reveals cellular
components of the phenotype}
How does a phenotype emerge? We realized that with the tools that we have
developed, it is possible to understand what tissues contribute to a phenotype
in a probabilistic framework. In other words, we can extract all genes
associated with a particular phenotype, then search for tissue terms that are
enriched to understand how a phenotype arises from interactions between
anatomical regions. As a test of this, we selected the egg-laying defective
(Egl) phenotype. In \cel{}, egg-laying is a complex behavior that involves a
large number of tissues~\cite{Li1990}. The somatic gonad acts as a repository
for the eggs, the uterine seam cells help protect the uterus, and a variety of
muscles help contract the uterus and open the vulva to lay an
egg~\cite{Sulston1977}. The vulva must be well-formed to allow passage of an
egg, and the hermaphrodite-specific neuron (HSN) is involved in the egg-laying
control~\cite{Schafer2005}. The complexity of the interactions that happen to
allow egg-laying make understanding the Egl phenotype in terms of tissues a
challenging task.

We extracted all of the \cel{} genes that have been associated with an Egl
phenotype and we used TEA to understand what tissues are enriched. The HSN was
enriched more than five-fold above background (\qval{7}) as were vulD, vulC,
vulE and vulF (\qval{6}). The vulA, vulB2 and vulB1 were enriched at slightly
lower levels (\qval{5}), whereas the uterine muscles and uterine seam cells were
enriched more than twice above background levels (\qval{2}) (see
Fig.~\ref{fig:egl}). Therefore, the Egl phenotype would seem to emerge primarily
from defects in the HSN, secondarily from defects in the vulva, and only
sometimes from defects in the uterine seam cells or muscles. It is notable that
all vulval cells were not equally enriched. Although all the `vul' cells are
annotated to a similar degree (between 50--70 genes for each cell type), the
vulD and vulC cells had the largest enrichment effect size and the lowest
q-values, suggesting that these cells are more likely to be associated with an
Egl phenotype than the others. This may reflect the fact that vulD and vulE are
the site of attachment for four vulval muscles, vm1. Perhaps this attachment is
particularly fragile, and perturbations to these cells prevent adequate function
of these muscles. In support of these observations, P7.pa had the largest
fold-enrichment of any tissue. In \cel{}, P7.pa gives rise to vulD and vulC.
However, vulF is also attached to a set of four additional vulval muscles, vm2.
Why is vulF less associated with an Egl phenotype?

\begin{figure}[htbp]
  \renewcommand{\familydefault}{\sfdefault}\normalfont{}
  \centering
  \includegraphics[width=.6\linewidth]{Egl-phenotype.pdf}
  \caption{The Egl phenotype is a complex phenotype that is the result of
  interactions between many tissues. To dissect the contributions of individual
  tissues to generating an Egl phenotype, we obtained all genes annotated with
  with an Egl term. \textbf{A} We used TEA on the set of Egl genes to identify
  enriched tissues. \textbf{B} Anatomic diagram showing the tissues that are
  most enriched in the Egl gene set. Color coding shows the qualitative
  ordering of enrichment (red-Most enriched, yellow-least enriched). For clarity,
  not all cells are shown. All vm1 (4 cells) and vm2 (4 cells) muscles are
  symmetrical arranged around the vulva, but only 2 cells are shown for each.
  There are two seam cells on the left and right side of the vulva, but only
  cell on the right is shown. Only terminally differentiated cells are shown.}
\label{fig:egl}
\end{figure}

\subsubsection*{Quantifying the anatomy-phenotype
                mapping via Bayesian probabilities}
Another way to understand the phenotype-anatomy mapping is by considering how
informative a given anatomy term is on a particular phenotype, or vice-versa. To
this end, we calculated two conditional probabilities that helped us answer this
question. The first conditional probability,
\begin{equation}
  P( \text{a gene has Egl annotation} | \text{it is expressed in } X)
\end{equation}
answers the question: For a gene with an expression pattern that includes the
tissue term X (i.e., the gene is expressed at least in X), what is the
probability that this gene has an Egl phenotype (i.e., the phenotype annotations
for this gene include Egl)? For simplicity, we can re-write this equation more
succintly by removing a few words. The calculation of this probability is
straightforward and follows from the definition of conditional probability:
\begin{equation}
  P(\text{Egl}|X) = \frac{N_\text{genes annotated Egl and X}}
                         {N_\text{genes annotated with X}}.
\label{egl_x}
\end{equation}
Equation~\ref{egl_x} measures how likely a gene is to be annotated with an Egl
phenotype given that its expression pattern includes the term $X$. A related
quantity (which is neither the inverse nor the complement) is the conditional
probability that a gene which is annotated with at least the Egl phenotype is
expressed in tissue $X$. That is to say,
\begin{equation}
  P(X|\text{Egl}) = \frac{N_\text{genes annotated Egl and X}}
                         {N_\text{genes annotated with Egl}}.
  \label{x_egl}
\end{equation}
Equation~\ref{x_egl} tells us how probable it is that any given gene that is
annotated with an Egl phenotype includes $X$ as a tissue term. Taken together,
equations~\ref{egl_x} and~\ref{x_egl} help us understand how predictive anatomic
expression is of phenotypes, and how predictive phenotypes are of anatomic
expression.

\begin{table*}
  \renewcommand{\familydefault}{\sfdefault}\normalfont{}
  \centering{}
  \ra{1.3}
  \begin{tabular}{@{}lcc@{}}
  \toprule{}Tissue & $P(\text{Egl}| X)$ & $P(X|\text{Egl})$\\
  \bottomrule{}P7.pa & $0.30$ & $0.04$\\
  HSN & $0.27$ & $0.11$\\
  vulC & $0.27$ & $0.06$\\
  vulD & $0.26$ & $0.07$\\
  vulE & $0.25$ & $0.06$\\
  vulF & $0.24$ & $0.06$\\
  vulA & $0.24$ & $0.05$\\
  vulB2 & $0.23$ & $0.05$\\
  vulB1 & $0.22$ & $0.05$\\
  nervous system & $0.02$ & $0.72$ \\
  pharynx  & $0.00$ & $0.46$\\
  sex organ & $0.07$ & $0.41$\\
  tail & $0.05$ & $0.33$\\
  \bottomrule{}
  \end{tabular}
  \caption{Conditional probabilities for various tissues. The
  first column shows the conditional probability that a gene has an Egl
  phenotype given that it has expression in tissue $X$ (given by the row).
  The second column shows the conditional probability that a gene has expression
  in the anatomy term X given that it has an Egl phenotype. The first 9 terms
  are the terms for which $P(\text{Egl}|X)$ is maximized. The last three terms
  are the terms which have the highest $P(X|\text{Egl})$. For clarity, the Pn.p
  cells are not shown even though $P(\text{Egl}|\text{Pn.p})\sim 0.24$.}
\label{tab:cond_probs}
\end{table*}

We calculated the conditional probability that a gene has an Egl phenotype given
that it's expression pattern includes a tissue term $X$ and we searched for the
tissue terms that maximized this probability. The list of terms that maximized
this probability reflected the results from running TEA on the subset of genes
that have an Egl phenotype. We also calculated the conditional probability that
a gene has expression in a tissue term $X$ given that it is annotated with an
Egl phenotype and we searched for terms that maximized this probability (see
Table~\ref{tab:cond_probs}). The terms that maximized this probability were
`nervous system', `pharynx' (a body part with a lot of neurons), `sex organ' and
`tail' (a body part with neurons and hypodermis). In general, the terms that had
a high $P(\text{Egl}|X)$ did not have a high $P(X|\text{Egl})$. Additionally,
the terms that had a high $P(X|\text{Egl})$ are broad terms that include a lot
of cells, whereas the terms that had a high $P(\text{Egl}|X)$ were considerably
more specific. We conclude that the Egl phenotype arises from a small set of
tissues. The Egl phenotype can be best predicted by genes with expression
patterns that include at least one of a small number of cells (mainly vul cells,
HSN). On the other hand, answering whether the expression pattern of a gene
includes a particular anatomic region or tissue given that the gene has an Egl
phenotype is hard to do for small tissues or single cells. However, guesses
about what functional system or broad anatomic region an Egl gene is expressed
in can be answered with confidence ($\sim70\%$ of the time, an Egl-causing gene
is expressed in the nervous system).

\section*{Conclusions}
The addition of GO and Phenotype Ontology enrichment testing to WormBase marks
an important step towards a unified set of analyses that can help researchers to
understand genomic datasets. These enrichment analyses will allow the community
to fully benefit from the data curation ongoing at WormBase. By using the same
algorithms to generate enrichment dictionaries for testing and using the same
model to test for term enrichment, these tools provide a coherent framework with
which to analyze genomic data. In particular, it is our hope that phenotype
enrichment will be of use to geneticists performing genome-wide analyses,
because they are familiar with the ontological terms that are tested and with
their biological meaning. Ideally, PEA could allow researchers to design better
screens to maximize target gene identification by quantifying the phenotypes
that are most overrepresented. An intriguing new direction of research would be
to create a controlled language for screening methods. With such a language, we
should be able to computationally suggest to a researcher what screens one may
wish to perform given a dataset.

\bibliography{pea-citations}
\bibliographystyle{pnas2011}

\end{document}